An efficient acyclic contact planner for multiped robots

### 2017-01-12 body/results.tex
updated figure on julien s advice, integrated [CONTACT] modifs by Andrea

At the end of this Section, we analyze the influence of the parameters of our framework on the output, and provide the \textit{interactive} computation times obtained in each case.
We also compare the times obtained with HRP-2 with respect to previous works.

@@

\subsubsection{On the choice of heuristics:} \label{sec:heuristichoices}
In our conference paper, the computed motions were only generated using the EFORT heuristic.
EFORT is designed for tasks requiring to exert important forces (such as pushing / pulling / climbing). 
Regarding locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this does not significantly impact
the generation of the plan, the resulting interpolation on the real robot turned out to be harder \citep{Carpentier2016}.
For this reason, we prefer to use our manipulability-based heuristic for the legs of the robot, but we still
use EFORT for the arms, that results in less contact repositioning.

@@

\begin{table*}
\centering
\footnotesize
\begin{tabular}{ >{\centering\arraybackslash}m{37pt} | >{\centering\arraybackslash}m{57pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{70pt} | >{\centering\arraybackslash}m{73pt} | >{\centering\arraybackslash}m{80pt} | >{\centering\arraybackslash}m{10pt}}
  Scenario (nb steps) &  Complete guide generation (ms) & Static equilibrium (ms) & Collision (ms) & Inverse Kinematics (ms) & Total generation time (ms) & Time per step (ms)\\
 \hline
   Stairs (18) & 5 -- \textbf{6} --  18 & 13 --  \textbf{32} -- 329   & 1 --  \textbf{4} -- 38 & 26 --  \textbf{127} -- 1345 & 92 --  \textbf{261} -- 2174 & \textbf{15} \\
   %~ Standing (24)& 4 -- \textbf{451} --  985 & 27 --  \textbf{144} -- 338   & 2 --  \textbf{12} -- 37 & 144 --  \textbf{1046} -- 2374 & 310 --  \textbf{1622} -- 3429 & \textbf{66}  \\
   Standing (24)& 65 -- \textbf{1086} --  5227 & 27 --  \textbf{144} -- 338   & 2 --  \textbf{12} -- 37 & 144 --  \textbf{1046} -- 2374 & 371 --  \textbf{2257} -- 7671 & \textbf{94}  \\
   %~ Car (61)& 1 -- \textbf{529} --  995 & 64 --  \textbf{85} -- 144   & 394 --  \textbf{735} -- 1422 & 3947 --  \textbf{13069} -- 36262 & 6775 --  \textbf{20890} -- 52963 & \textbf{342} \\
   Car (61)& 145 -- \textbf{1055} --  15063 & 64 --  \textbf{85} -- 144   & 394 --  \textbf{735} -- 1422 & 3947 --  \textbf{13069} -- 36262 & 6919 --  \textbf{21416} -- 67031 & \textbf{351} \\
   %~ Rubble (71)& 3 -- \textbf{524} --  993 & 242 --  \textbf{511} -- 3480   & 233 --  \textbf{505} -- 4564 & 180 --  \textbf{414} -- 3518 & 1400 --  \textbf{3058} -- 13126 & \textbf{43} \\
   Rubble (71)& 151 -- \textbf{15220} --  125867 & 242 --  \textbf{511} -- 3480   & 233 --  \textbf{505} -- 4564 & 180 --  \textbf{414} -- 3518 & 1548 --  \textbf{18058} -- 13126 & \textbf{254} \\
   %~ Race (112)& 1 -- \textbf{501} --  995 & 266 --  \textbf{449} -- 3956   & 824 --  \textbf{1061} -- 2130 & 666 --  \textbf{874} -- 1613 & 2530 --  \textbf{3722} -- 10030 & \textbf{33}
   Race (112)& 267 -- \textbf{966} --  2257 & 266 --  \textbf{449} -- 3956   & 824 --  \textbf{1061} -- 2130 & 666 --  \textbf{874} -- 1613 & 2797 --  \textbf{4187} -- 11292 & \textbf{37}
   %Standing up & $70$ & 23 & 48\\
 \end{tabular}
\caption{minimum, \textbf{average} and worst time (in ms) spent in the generation process for each scenario and each critical part of the generation process (not all parts are timed,
thus the average total computation time is higher than the sum of each part). The last
column indicates the average time necessary to compute one contact transition. }
\label{tab:requestime}
\quad
\end{table*}

\begin{table*}
\centering
\begin{tabular}{ l | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | c}
  &  Path planning & Equilibrium feasibility & Kinematic failure & Equilibrium failure \\
 \hline
   Steep stairs & 100\%  & 99.5\% & 0.1\% & 0.4\% \\
   Standing up & 68\% & 87.8\% & 6.1\% & 6.1\% \\
   Car egress & 39\% & 77.0\% & 21.0\% &  2.0\% \\
   Rubble & 74\% & 97.9\% & 0.1\% & 2.0\% \\
   Obstacle race & 58.0\% & 95.7\% & 1.8\% & 2.5\% \\
 \end{tabular}
\caption{Success rates for each scenario, rounded to the first decimal. The first value indicates the percentage of successful complete contact plannings for 1000 tests; The second value
indicates the percentage of equilibrium feasible root configurations: considering each limb individually, indicates the percentage of root configurations of the guide path that
led to a feasible contact. Kinematic failure is the percentage of contact generations that failed because no collision-free candidate was found. Equilibrium failure is the percentage of contact
generations that failed because no candidate respected the static equilibrium condition.}
\label{tab:requestpercent}
\quad
\end{table*}

@@

Considering the repartition of the computation time, for HRP-2, most of the time is spent performing inverse kinematics.
This is not surprising considering the number of calls to the methods: IK projection is used intensively to maintain contact continuity between two postures; 
it is also applied every time a new candidate needs to be evaluated. In particular for the car egress scenario,
the kinematic constraints are very demanding. Therefore a large number of projections fail because of joint-limit violations.

@@

In all scenarios, one can observe that the average computation time for one single step is largely below one second,
thus allowing to consider \gls{interactive} applications.

\subsubsection{Success rates:}
Table~\ref{tab:requestpercent} summarizes the success rates obtained for each scenario.
As expected, our planner does not succeed systematically, because of the approximations made in our formulation. The extreme situation of the car egress scenario provides the less satisfying results in terms of kinematic failure. This is not
surprising considering the narrowness of the scene.
\deladp{From a pragmatic point of view, }regarding \deladp{the} computation times, \deladp{we claim that} our approach provides a satisfying compromise between completeness and efficiency.
Indeed, \deladp{the advantage of the framework is that} when the contact generation fails, it does so rapidly, which allows us to rapidly re-plan \deladp{a new contact sequence} with a reasonable chance of success.
The most efficient (and immediate) approach is probably to launch in parallel several instances of the planner \deladp{for a given problem} (our current implementation is single threaded) and to apply any successful result.

\begin{table*}
\centering
\begin{tabular}{ c | c | c | c | c }
 Scenario & Method  & Guide Path & Contact sequence & Interpolation \\
 \hline
   \multirow{3}{*}{Stair 20 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}5.42 min}  \\
							 & \textbf{Ours} + \citeauthor{Carpentier2016}  & \textbf{5 ms} & $\mathbf{<}$ \textbf{18 ms} & $ <$ 2s \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{3}{*}{Stair 30 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}4.08 min}  \\
							 & \textbf{Ours}  & \textbf{5 ms} & $\mathbf{<}$ \textbf{18 ms} & X \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{3}{*}{Stair 40 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}10.08 min}  \\
							 & \textbf{Ours}  & \textbf{5 s} & \textbf{3 s} & X \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{2}{*}{Table (car) egress} & \citeauthor{Bouyarmane2009, DBLP:conf/iser/EscandeKMG08} & 10 min & \multicolumn{2}{c}{\cellcolor{blue!25}3.5 hours}  \\
							 & \textbf{Ours}  & \textbf{529 ms} & \textbf{$<$ 22 s} & X \\
 \end{tabular}
\caption{Comparison between the computation times obtained by our method and previous ones, for each phase of the planning. Blue cells indicate merged times (because no
precise time information is available). The X symbol indicates scenarios where we did not address the interpolation of the contact sequence.}
\label{tab:compprev}
\quad
\end{table*}
\subsection{Comparison with previous work} \label{sec:compa}
Comparing our method with others is not trivial. Indeed, we focus on the contact planning aspect of the problem, and do not address
the interpolation in this paper, nor provide a strong guarantee that our sequences can be interpolated.
Most papers in the literature only provide the overall computation
time, without specifying how much is spent in each phase. Moreover, there exists no off-the-shelf implementation
of the related methods. Due to their complexity, we cannot afford the development resources required to reimplement them.

Furthermore, the interpolation itself is not completely solved. For instance in \cite{DBLP:conf/iser/EscandeKMG08}, the end-effector trajectories are computed automatically with splines, but are manually corrected if they enter in collision. In \citep{Mordatch:2012:DCB:2185520.2185539}, internal joint collisions can occur in the resulting motion.
Recent methods suffer from similar issues \citep{Carpentier2016}.

>>>>

At the end of this Section, we discuss the role of the parameters of our framework on the output. We then provide the \textit{interactive} computation times obtained in each case.
We also compare the times obtained with HRP-2 with respect to previous works.

@@

\subsubsection{On the choice of heuristics:} \label{sec:heuristichoices}
In our conference paper, the computed motions were only generated using the EFORT heuristic.
EFORT is designed for tasks requiring to exert important forces (such as pushing / pulling / climbing). 
Regarding locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this does not significantly impact
the generation of the plan, the resulting interpolation on the real robot turned out to be harder.
For this reason, we prefer to use our manipulability-based heuristic for the legs of the robot, but we still
use EFORT for the arms, that results in less contact repositioning.

@@

\begin{table*}
\centering
\footnotesize
\begin{tabular}{ >{\centering\arraybackslash}m{37pt} | >{\centering\arraybackslash}m{57pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{70pt} | >{\centering\arraybackslash}m{73pt} | >{\centering\arraybackslash}m{80pt} | >{\centering\arraybackslash}m{10pt}}
  Scenario (nb steps) &  Complete guide generation (ms) & Static equilibrium (ms) & Collision (ms) & Inverse Kinematics (ms) & Total generation time (ms) & Time per step (ms)\\
 \hline
   Stairs (18) & 5 -- \textbf{6} --  18 & 13 --  \textbf{32} -- 329   & 1 --  \textbf{4} -- 38 & 26 --  \textbf{127} -- 1345 & 92 --  \textbf{261} -- 2174 & \textbf{15} \\
   %~ Standing (24)& 4 -- \textbf{451} --  985 & 27 --  \textbf{144} -- 338   & 2 --  \textbf{12} -- 37 & 144 --  \textbf{1046} -- 2374 & 310 --  \textbf{1622} -- 3429 & \textbf{66}  \\
   Standing (24)& 65 -- \textbf{1086} --  5227 & 27 --  \textbf{144} -- 338   & 2 --  \textbf{12} -- 37 & 144 --  \textbf{1046} -- 2374 & 371 --  \textbf{2257} -- 7671 & \textbf{94}  \\
   %~ Car (61)& 1 -- \textbf{529} --  995 & 64 --  \textbf{85} -- 144   & 394 --  \textbf{735} -- 1422 & 3947 --  \textbf{13069} -- 36262 & 6775 --  \textbf{20890} -- 52963 & \textbf{342} \\
   Car TODO (61)& 145 -- \textbf{1055} --  15063 & 64 --  \textbf{85} -- 144   & 394 --  \textbf{735} -- 1422 & 3947 --  \textbf{13069} -- 36262 & 6919 --  \textbf{21416} -- 67031 & \textbf{351} \\
   %~ Rubble (71)& 3 -- \textbf{524} --  993 & 242 --  \textbf{511} -- 3480   & 233 --  \textbf{505} -- 4564 & 180 --  \textbf{414} -- 3518 & 1400 --  \textbf{3058} -- 13126 & \textbf{43} \\
   Rubble (71)& 151 -- \textbf{15220} --  125867 & 242 --  \textbf{511} -- 3480   & 233 --  \textbf{505} -- 4564 & 180 --  \textbf{414} -- 3518 & 1548 --  \textbf{18058} -- 13126 & \textbf{254} \\
   %~ Race (112)& 1 -- \textbf{501} --  995 & 266 --  \textbf{449} -- 3956   & 824 --  \textbf{1061} -- 2130 & 666 --  \textbf{874} -- 1613 & 2530 --  \textbf{3722} -- 10030 & \textbf{33}
   Race (112)& 267 -- \textbf{966} --  2257 & 266 --  \textbf{449} -- 3956   & 824 --  \textbf{1061} -- 2130 & 666 --  \textbf{874} -- 1613 & 2797 --  \textbf{4187} -- 11292 & \textbf{37}
   %Standing up & $70$ & 23 & 48\\
 \end{tabular}
\caption{minimum, \textbf{average} and worst time (in ms) spent in the generation process for each scenario and each critical part of the generation process (not all parts are timed,
thus the average total computation time is higher than the sum of each part). The last
column indicates the average time necessary to compute one contact transition. }
\label{tab:requestime}
\quad
\end{table*}

\begin{table*}
\centering
\begin{tabular}{ l | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | c}
  &  Path planning success rate & Success rate & Kinematic failure & Equilibrium failure \\
 \hline
   Steep stairs & 100\%  & 99.5\% & 0.1\% & 0.4\% \\
   Standing up & 68\% & 87.8\% & 6.1\% & 6.1\% \\
   Car egress TODO & 39\% & 77.0\% & 21.0\% &  2.0\% \\
   Rubble & 74\% & 97.9\% & 0.1\% & 2.0\% \\
   Obstacle race TODO & 58.0\% & 95.7\% & 1.8\% & 2.5\% \\
 \end{tabular}
\caption{ TODO AMBIGUITE POURCENTAGE Success rates for each scenario, rounded to the first decimal. The first value indicates the percentage of successful complete contact plannings for 1000 tests; The second value
indicates the percentage of equilibrium feasible root configurations: considering each limb individually, indicates the percentage of root configurations of the guide path that
led to a feasible contact. Kinematic failure is the percentage of contact generations that failed because no collision-free candidate was found. Equilibrium failure is the percentage of contact
generations that failed because no candidate respected the static equilibrium condition.}
\label{tab:requestpercent}
\quad
\end{table*}

@@

Considering the repartition of the computation time, for HRP-2, most of the time is spent performing inverse kinematics.
This is not surprising considering the number of calls to the methods: IK projection is used intensively to maintain contact continuity between two postures; 
it is also applied every time a new candidate needs to be evaluated. In particular for the car egress scenario,
the kinematic constraints are very demanding to avoid collisions.

@@

In all scenarios, one can observe that the average computation time for one single step is largely below one second,
thus allowing to consider \gls{interactive} applications. This would allow to consider online autonomous planning of the robot motion.

\subsubsection{Success rates:}
Table~\ref{tab:requestpercent} summarizes the success rates obtained for each scenario.
Despite the complexity of the scenarios adressed, as well as the approximations made in our formulation, our planner succeeds in a large majority of cases.
As expected, our planner does not succeed systematically, because of the approximations made in our formulation. The extreme situation of the car egress scenario provides the less satisfying results in terms of kinematic failure. This is not
surprising considering the narrowness of the scene.
Regarding computation times, our approach provides a satisfying compromise between completeness and efficiency.
Indeed, when the contact generation fails, it does so rapidly, which allows us to rapidly re-plan with a reasonable chance of success.
The most efficient (and immediate) approach is probably to launch in parallel several instances of the planner (our current implementation is single threaded) and to apply any successful result.

%~ \begin{table*}
%~ \centering
%~ \begin{tabular}{ c | c | c | c | c }
 %~ Scenario & Method  & Guide Path & Contact sequence & Interpolation \\
 %~ \hline
   %~ \multirow{3}{*}{Stair 20 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}5.42 min}  \\
							 %~ & \textbf{Ours} + \citeauthor{Carpentier2016}  & \textbf{5 ms} & $\mathbf{<}$ \textbf{18 ms} & $ <$ 2s \\
							 %~ & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 %~ \hline
   %~ \multirow{3}{*}{Stair 30 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}4.08 min}  \\
							 %~ & \textbf{Ours}  & \textbf{5 ms} & $\mathbf{<}$ \textbf{18 ms} & X \\
							 %~ & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 %~ \hline
   %~ \multirow{3}{*}{Stair 40 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}10.08 min}  \\
							 %~ & \textbf{Ours}  & \textbf{5 s} & \textbf{3 s} & X \\
							 %~ & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 %~ \hline
   %~ \multirow{2}{*}{Table (car) egress} & \citeauthor{Bouyarmane2009, DBLP:conf/iser/EscandeKMG08} & 10 min & \multicolumn{2}{c}{\cellcolor{blue!25}3.5 hours}  \\
							 %~ & \textbf{Ours}  & \textbf{529 ms} & \textbf{$<$ 22 s} & X \\
 %~ \end{tabular}
%~ \caption{Comparison between the computation times obtained by our method and previous ones, for each phase of the planning. Blue cells indicate merged times (because no
%~ precise time information is available). The X symbol indicates scenarios where we did not address the interpolation of the contact sequence.}
%~ \label{tab:compprev}
%~ \quad
%~ \end{table*}

\begin{table*}
\centering
\begin{tabular}{ c | c | c | c | c }
 Scenario & Method  & Guide Path & Contact sequence & Interpolation \\
 \hline
   \multirow{3}{*}{Stair 20 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}5.42 min}  \\
							 & \textbf{Ours} + \citeauthor{Carpentier2016}  & \textbf{5 ms} & $\mathbf{<}$ \textbf{18 ms} & $ <$ 2s \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{3}{*}{Stair 30 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}4.08 min}  \\
							 & \textbf{Ours}  & \textbf{5 ms} & $\mathbf{<}$ \textbf{18 ms} & X \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{3}{*}{Stair 40 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}10.08 min}  \\
							 & \textbf{Ours}  & \textbf{5 s} & \textbf{3 s} & X \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 %~ \hline
   %~ \multirow{2}{*}{Table (car) egress} & \citeauthor{Bouyarmane2009, DBLP:conf/iser/EscandeKMG08} & 10 min & \multicolumn{2}{c}{\cellcolor{blue!25}3.5 hours}  \\
							 %~ & \textbf{Ours}  & \textbf{529 ms} & \textbf{$<$ 22 s} & X \\
 \end{tabular}
\caption{Comparison between the computation times obtained by our method and previous ones, for each phase of the planning. Blue cells indicate merged times (because no
precise time information is available). The X symbol indicates scenarios where we did not address the interpolation of the contact sequence.}
\label{tab:compprev}
\quad
\end{table*}

\subsection{Validation of the contact plans, and comparison with previous work} \label{sec:compa}
To validate our plans, we had to implement our own interpolation method to address $\mathcal{P}_3$.
We refer the curious reader to \citep{Carpentier2016} for the implementation details of our method. 
This motion is in turn validated in our dynamic simulator~\citep{DelPrete2015b}. The stair climbing scenario was validated on the real HRP-2 robot \citep{Carpentier2016}.
The obtained motions are available in the video.

Few contributions provided computation times on the complete problem, necessary to provide comparison between our approaches. 
Furthermore, $\mathcal{P}_3$ is an open problem in the presence of obstacles. For instance in \cite{DBLP:conf/iser/EscandeKMG08}, the end-effector trajectories are computed automatically with splines, but are manually corrected if they enter in collision. In \citep{Mordatch:2012:DCB:2185520.2185539}, internal joint collisions can occur in the resulting motion.
Recent methods suffer from similar issues \citep{Carpentier2016}.
The only valid scenarios addressed completely and automatically in previous works are thus the stair climbing scenarios with HRP-2 
To compare our approach with previous contributions, we considered those that addressed the complete problem, and selected scenarios similar to ours, thus focusing on the HRP-2 robot.

Table~\ref{tab:compprev} presents the results we obtained
